### 2011-03-25 workflow.Rnw
More work

\documentclass[12pt]{article}
\usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs,listings}
\usepackage[colorlinks=TRUE]{hyperref}
\usepackage{microtype}

\title{Six steps to a better relationship with your future self.}

\author{Jake Bowers\thanks{I owe many thanks to Mark Fredrickson,
    Brian Gaines, Kieran Healy, Kevin Quinn and Cara
    Wong for direct help on this document and to Mika LaVaque-Manty
    and Ben Hansen for many useful discussions on this topic. The
    source code for this document may be freely downloaded and
    modified from \url{https://github.com/jwbowers/workflow}.}}

\date{\today}

\newcolumntype{.}{D{.}{.}{1.2}}

\usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize}
\DefineVerbatimEnvironment{Scode}{Verbatim}  {xleftmargin=.5em,fontsize=\footnotesize}
\fvset{listparameters={\setlength{\topsep}{0pt}}}
\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}
\SweaveOpts{keep.source=TRUE}  %Show comments
\fvset{fontsize=\footnotesize} %slightly smaller verbatim font

%% These next lines tell latex that it is ok to have a single graphic
%% taking up most of a page, and they also decrease the space arou
%% figures and tables.
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}
\setlength{\intextsep}{2ex}
\setlength{\floatsep}{2ex}
\setlength{\textfloatsep}{2ex}

\begin{document}
\frenchspacing
\VerbatimFootnotes
\maketitle

<<initialize,echo=FALSE,results=hide>>=
  ##First, just setup the \R environment:
  options(SweaveHooks=list(
            fig=function(){
              par(mar=c(3.5, 3, 1.1, 0),
                  pty="s",
                  mgp=c(1.5,0.5,0),
                  oma=c(0,0,0,0))},#,
            echo=function(){options(continue=" ") ##Don't show "+" prompts,
                            options(prompt=" ")}
            ),
          digits=3,
          scipen=8,
          width=50,
          error=function(){options(prompt="> ",continue="+ ");NULL}
          )

@

\begin{verse}
Do I contradict myself? \\
Very well then I contradict myself, \\
(I am large, I contain multitudes.) \\
\citep{whitman1855song}
\end{verse}

An idea is born in a coffee shop, a seminar, a quiet walk. On this
gray day in 2011, the idea dispells February's doldrums. The student
rushes home, mind racing, the cold ignored.
% Now he knows why he went to graduate school.

This idea inspires a seminar paper in the spring. A conference paper
arises from the seminar paper in collaboration with another student in
2012. A dissertation chapter descends from the conference paper in
2013. Other dissertation chapters take up 2014. A submission to a
journal with the original co-author and a new collaborator happens in
2015. Revision and resubmission wait until 2017 while harried editors,
reviewers and authors strive to balance research, teaching, service,
and life. By now, the three lucky collaborators work as professors in
three different universities. In 2018 a child is born and a paper is
published. The United Nations takes an interest in the paper in 2019
and hosts a conference to discuss implications of the research. In
2020 a first year graduate student in a coffee shop has an idea that
challenges the results in the now famous paper. What would happen if
the authors had controlled for X? Or included information now
available but missing in 2012?  Or chosen a different likelihood
function?  Will the United Nations (now eager to act based the paper)
make a wrong move?

The first author convenes a three way video conference with the other
collaborators during his homeward commute after putting his flying car
in auto-drive mode.\footnote{One assumes that video chatting during
  manual driving of flying cars will have been outlawed in his state
  by 2020.} The group must go back to the analyses. Which ones?  The
ones from 2011?  Or 2012? Or 2018? Where are the files? The next day,
one member of the group who has kept some hard-drives around out of
nostalgia finds some of the files.\footnote{This is the same guy who
  still owns cassette tapes and compact discs.} Now
re-analyses should be easy. Right? The student, now professor, should
remember the reason for those bits of code (or at least should
remember which series of mouse clicks were used to produce the numbers
for that crucial table --- and that mousing should not be time
consuming to redo exactly as it was done in 2011 \ldots or was it
2015?). Right? And, of course, the way Microsoft
Word/Stata/SPSS/R/LISREL understands files and the way that 
machines in 2020 read and write them is the same --- since Windows and
Mac OS X have always existed and will always continue to exist more or
less as they currently exist? Right? And the group knows exactly which
bit of code produced which table and which figure, right? And they
wrote their code following Nagler's Maxims \citep{nagler1995coding}
and King's Replication Standard \citep{king1995replication} right?

Right? Wrong? If the collaborators find themselves say ``Wrong'' in
answer to the questions posed here then reproducing, updating, or
changing the original analyses will take a lot of time. If
reproduction is hard to do, then the reputations of the scholars will
suffer and, more importantly, world peace will have been delayed. This
essay provides some suggestions for practices which will make such
reproduction occur much more easily and quickly in the event that
famous papers require special scrutiny. Specifically, this piece aims
to amplify some of what we already ought to know
\citep{king1995replication,nagler1995coding}, and to add to some of
those ideas given current practices, platforms, and possibilities.

%%% This data frame will be used throughout the analyses.
%%% Putting it up top, so we can use it in the comments section.
<<read.data,echo=FALSE,results=hide>>=
##  This read.csv line is slightly more complex than needed because of
##  the particular encoding used for characters in the source file.
dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
good.df$gini04<-good.df$Gini2004/100
good.df$protac00<-good.df$protac2000
@

\section{Data analysis is computer programming.}\label{sec:data-analys-comp}

All results (numbers, comparisons, tables, figures) should arise from
code, not series of mouse clicks or copying and pasting. If I wanted
to re-create the figure you created but including a new variable or
specification, I should be able to do so with just a few edits to the
code rather than knowledge of how you used your pointing device in
your graphical user interface.

Using R \citep{R:2011}, for example, I might specify that the file \Verb+fig1.pdf+
was produced by the following commands in a file called
\Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some other
  R functions used in this essay come from the
  \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
  data analysis. Functions like \Verb+please-plot+ can
  be blocked and more polite versions such as \Verb+may-I-please-have-a-plot+ can be required using
  \Verb+options(politeness=99)+ }

\begin{Verbatim}[fontsize=\footnotesize]
thedata <- read.csv("Data/thedata-15-03-2011.csv") ## Read the data
pdf('fig1.pdf') ## begin writing to the pdf file
please-plot(outcome by explanatory using thedata. red lines please.)
please-add-a-line(using model1)
## Note to self: a quadratic term does not add to the substance
## model2<-please-fit(outcome by explanatory+explanatory^2 using thedata
## summary(abs(fitted(model1)-fitted(model2)))
dev.off() ## stop writing to the pdf file
\end{Verbatim}

Now, in the future if I wonder how ``that plot on page 10'' was
created, I will know: (1) ``that plot'' is from a file called
\Verb+fig1.pdf+ and (2) \Verb+fig1.pdf+ was created in
\Verb+fig1.R+. In a future where R still exists, changing the figure
will require quick edits of commands already written.  In a future
where R does not exist, I will at least be able to read the plain text
R commands and use them to write code in my new favorite statistical
computing language: R scripts are written in
\href{http://en.wikipedia.org/wiki/Plain_text}{plain text}, and plain
text is a format that will be around as long as computer programmers
write computer programs. \footnote{What is more, since R is open
  source, I will be able to download an old version of R, download an
  old fashioned open-source operating system (like Ubuntu 10), and
  run the old-fashioned statistical computing environment in the
  old-fashioned operating system in a virtual machine on my
  new-fashioned actual machine.}

Moreover, realize that file names send messages to your future
self. Name your files with evocative and descriptive names. Your
collaborators are less likely to call you at midnight asking for help
if your files are named \Verb+inequality-and-protest-figures.R+ than
if your files are called \Verb+temp9+ or
\Verb+supercalifragilisticexpialidocious+. The use of the extension
\Verb+.R+ tells us and the operating system that the file contains R
commands. This part of the filename enables us to quickly search our
antique hard drives for files containing R scripts.

\paragraph{Step 1} Know the provenance of your results so that your
future self or current collaborators can quickly and easily reproduce
and thus change your work.% --- to know what
% commands created which figures and tables.

\section{No data analyst is an island for long.}
Data analysis involves a long series of decisions. Each decision
requires justification.  Some decisions will be too small and
technical for inclusion in the published article itself. These need to
be documented in the code itself \citep{nagler1995coding}. Paragraphs
and citations in the publication will justify the most important
decisions. So, one must code to communicate with yourself and
others. There are two main ways to avoid forgetting the reasons you
did something with data: comment your code and tightly link your code
with your writing.\footnote{One can also try the R command
  \Verb+put-it-in-my-mind(reason,importance='high')+ to firmly place a
  reason for a decision into the mind of the analyst. I myself have
  not had much luck with this function.}

\subsection{Code to communicate: Comment your code.}
Comments --- unexecuted text inside of a script --- are a message to
collaborators (including your future self) and other consumers of your
work. In the above code chunk, I used comments to explain the lines to
readers unfamiliar with R and to remember that I had tried a different
specification but decided not to use it because adding the squared
term did not really change the substantive story arising from the
model.\footnote{R considers text marked with \Verb+#+ as a comment.}
Messages left for your future self (or near-future others) help
retrace and justify your decisions as the work moves from seminar
paper to conference paper to poster back to paper to dissertation and
onwards.

% Luckily, if you are
% using a literate programming practice (ex. Sweave or odfWeave), you
% can write paragraphs to surround your code as well as technical
% comments in the code itself. If you are not using a strictly literate
% programming practice then use whatever commenting protocol exists in
% your analysis language. R, for example, considers text marked with
% ``\#'' as a comment. Here for example I comment about a linear model
% run for a paper and also include comments about another model run for
% due diligence.

% <<reg1,echo=TRUE,results=verb>>=
% ##  Repeat the regression run for Table~\ref{tab:protest}
% ##  Do political rights predict protest reporting?
% lm1 <- lm(protac00 ~ gini04 + meanpr, data=good.df)
% ##  Q: Was it worthwhile controlling for political rights?
% lm2 <- lm(protac00~gini04,data=good.df)
% coef(lm2)[2]-coef(lm1)[2]
% ##  A: Yes (around 3/4 an act of protest difference)
% @

Notice one another benefit of coding for an audience: we learn by
teaching. By assuming that others will look at your code, you will be
more likely to write clearer code, or perhaps even to learn more about
what you are doing as you do it.

Comment liberally. Comments are discarded when R runs analysis or
\LaTeX~turns plain text into page images, so only those who dig into
your work will see them.

\subsection{Code to communicate: Literate programming.}

Imagine you discover something new (or confirm something old). You
produce a nice little report on your work for use in discussions of
your working group or as a memo for a web or reviewer appendix. The
report itself is a pdf file or some other format which emphasizes
reading. Eventually pieces of that report (tables, graphs, paragraphs)
ought to contribute to the publishable paper. Re-create those analyses
by pointing, clicking, copying, or pasting would invite typing error
and waste time. Re-creating your arguments justifying why you dropped
China but not India from your analyses or why you used a quadratic
term but not a cubic term would also waste time. More importantly, we
and others want to know why we did what we did. Such explanations may
not be very clear if we have some pages of printed code in one hand
and a manuscript in the other.

How might one avoid these problems?
\href{http://en.wikipedia.org/wiki/Literate_programming}{Literate
  programming} is the practice of weaving code into a document ---
paragraphs, equations, and diagrams can explain the code, and the code
can numbers, figures, and tables (and diagrams and even equations and
paragraphs). Literate programming is not merely fancy commenting but
is about enabling the practice of programming itself to enable easy
reproduction and communication.

For example, in \S\ref{sec:data-analys-comp}, I suggested that we know
where ``that plot on page 10'' comes from by making sure we had a
\Verb+fig1.pdf+ file produced from a clearly commented plain text file
called something like \Verb+fig1.R+.  An even easier solution would be to
directly include a chunk of code to produce the figure inside of the
paper itself. This paper, for example, was written in plain text using
\LaTeX~ markup with R code chunks to make things like
Figure~\ref{fig:giniprot}: this combination of \LaTeX and R is called
Sweave \citep{Leis:2005}.\footnote{Support for Sweave is included with
  R.}

\begin{Verbatim}[fontsize=\footnotesize]
This paper, for example, was written in plain text using \LaTeX
markup with R code chunks to make things like
Figure~\ref{fig:giniprot}: this combination of \LaTeX and R is
called Sweave \citep{Leis:2005}.\footnote{Support for Sweave is
  included with R.}
\begin{figure}[H]
\centering
 <<fig1plot,echo=FALSE,fig=TRUE>>=
 ##  Make a scatterplot of Protest by Inequality
 with(good.df,plot(gini04,protac00,xlab='Gini Coefficient 2004 (UNDP)',
                   ylab='Mean Protest Activities\n(World Values Survey 1980-2000)'))
 ##  Label a few interesting points
 with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
      text(gini04,protac00,labels=Nation))
 @
\caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
\label{fig:giniprot}
\end{figure}
\end{Verbatim}

<<fig1code,echo=FALSE,fig=FALSE>>=
##  Make a scatterplot of Protest by Inequality
with(good.df,plot(gini04,protac00,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
                  cex=.8))

##  Label a few interesting points
with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
     text(gini04,protac00,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
@

\begin{figure}[H]
  \begin{center}
<<fig1plot,echo=FALSE,fig=TRUE,width=4,height=4>>=
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
<<fig1code>>
@
    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
    \label{fig:giniprot}
  \end{center}
\end{figure}

By using \Verb+\label{fig:giniprot}+, I do not need to keep track of
the figure number. Nor do I need a separate \Verb+fig1.R+ file or
\Verb+fig1.pdf+ file.  Tables and other numerical results are also
possible to generate within in the source code of a
scholarly paper. Those who view the code for this essay will see how
Table~\ref{tab:protest} was also generated directly from a regression
object.\footnote{\citet{beck2010reg} inspired this
  particular presentation of a linear model.}

<<maketable, echo=FALSE, results=hide>>=
  lm1<-lm(protac00~gini04+meanpr,data=good.df)  #  Run the regression

##  The next lines extract relevant and useful output from the
##  regression object, following Neal Beck's TPM article 2010.
themat<-cbind(summary(lm1)$coef[,c("Estimate","Std. Error")],confint(lm1))
n<-length(lm1$fitted.values)
r2<-signif(summary(lm1)$r.squared,2)
est.sigma<-signif(summary(lm1)$sigma,2)
colnames(themat)[3:4]<-c("~{}","~{}")
row.names(themat)<-c("Intercept","Income Inequality (lower=more equal)","Mean Political Rights (lower=more rights)")

library(xtable)
thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
                caption="People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, adjusting for average political
  rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest")
@

<<printtable,echo=FALSE,results=tex>>=
  print(thextab,include.colnames=FALSE,
        hline.after=NULL,
        table.placement="H",
        add.to.row=list(
          pos=list(0,3),
          command=c(
            "\\toprule \n &
             \\multicolumn{1}{c}{Coef} &
             \\multicolumn{1}{c}{Std. Err.} &
             \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
            paste("& \\multicolumn{4}{c}{n: ",n,",",
                  "   resid.sd: ",signif(est.sigma),
                  ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
            )
          ))
@

Literate data analysis is not the same as Sweave, even if Sweave is a
nice implementation.\footnote{The R project has a task view devoted to
  \href{http://cran.r-project.org/web/views/ReproducibleResearch.html}{reproducible
    research} listing many of the different approaches to literate
  programming for R.} If your workflow does not involve \LaTeX~ and R,
you can still implement some of the principles
here. \href{http://www.lyx.org/}{LyX} offers a WYSIWYG environment for
\LaTeX~ which supports Sweave. And the \Verb+odfWeave+ package in R
allows the use of OpenOffice documents in exactly the same
way.\footnote{A quick Google search of ``Sweave for Stata'' turned up
  lots of resources for literate programming with Stata.} Something
other than than Sweave will make our lives easier in the future. Then
we'll change.

\paragraph{Step \thesection}

\begin{quote}
  Let us change our traditional attitude to the construction of
  programs: Instead of imagining that our main task is to instruct a
  computer what to do, let us concentrate rather on explaining to
  human beings what we want a computer to do.
  \citep[][p. 97]{knuth1984literate}
\end{quote}

If we focus on explaining our data analysis to human beings, we will
do a better job with the data analysis itself: we will learn as we
focus on teaching, and we will avoid errors and wasting time as we
ensure that others (including our future selves) can retrace our
steps.

Keep in mind the distinction between the
``source code'' of a document (i.e. what computation was required to
produce it) and the visible, type-set page image. Page images are
great for reading, but not great for reproducing or collaborating. The
source code of any document exchanged by the group must be available
and executable.

In order to know that ``that plot on page 10'' is from a file called
``fig1.pdf'' you need to document it somehow. I can imagine a system
with very disciplined use of a word processor perhaps using special
auxiliary ``listoffigures.txt'' or ``MANIFEST.txt'' or ``README'' or
``Makefile'' files or something. Right now it is hard to beat Sweave
or some plain text combination of \LaTeX (or some other markup based
document formatting system with support for citations, etc... ) with
some plain text based statistical programming environment (like R or
Stata or Matlab, etc.) as a system for enhancing the collaboration of
research groups doing quantitative work in political
science. Something other than than Sweave will make our lives easier
in the future. Then we'll change.\footnote{For example, X uses the
  org-mode markup system in this issue.}

\paragraph{Step \thesection} Comments and other documentation enable you to
learn through teaching, code efficiently, and avoid duplicating your
effort. Literate programming also enables you to avoid the kinds of
errors which occur when you re-type numbers in to tables or
graphic-producing systems or even into a document itself. Finally, and
most importantly, a document that can be ``run'' to reproduce all of
the results is a document that can more effectively spur discussion
and learning and cumulation of research as people no longer need spend
weeks attempting to reproduce research on which they desire to build.

\url{http://en.wikipedia.org/wiki/Literate_programming}

\section{Meaningful code requires data.}

All files containing commands operating on data must refer to a data
file. A reference to a data file is a line of code the analysis
program will use to operate on (``load''/ ``open'' / ``get'' /
``use'') the data file. One should not have to edit this line on
different computers or platforms in order to execute this command.
Using R, for example, all analysis files should have
\Verb+load('thedata.rda')+ or
\Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
equivalent line in them, and \Verb+thedata.csv+ should be stored in
some place easy to find (like in the same directory as the file or
perhaps in \Verb+'Data/thedata.rda'+). Of course, it never hurts to
drop in a comment pointing to the data file.

Where should one store data files? A obvious solution is always to
make sure that the data file used by a command file is in the same
directory as the command file. More elegant solutions require all
co-authors to have the same directory structure so that
\Verb+load('Data/thedata.rda')+ means the same thing on all computers
used to work on the project.

The principle of modularity suggests that you separate data cleaning,
processing, recoding, and merging from analysis in different files
\citep{nagler1995coding}. So, perhaps your analysis oriented files
will \Verb+load('cleandata.rda')+ and a comment in the code will alert
the future you (among others) that \Verb+cleandata.rda+ was created
from \Verb+create-cleandata.R+ which in turn begins with
\Verb+read.csv(\url('http://www.greatfreedata.gov/dirtydata.csv'))+. Such
a data processing file will typically end with something like
\Verb+save('cleandata.rda')+ so that we are doubly certain about the
provenance of the data.

Now, if in the future we wonder where \Verb+cleandata.rda+ came from, we
might search for occurrences of 'cleandata' in the files our
system. However, if such searching among files is a burden, an even
nicer solution is to maintain a file for each project called
``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists the
data and command files with brief descriptions of their functions and
relations.

\paragraph{Step \thesection} We should know data where the data came from and
what operations were performed on which set of data.

Let me note that this principle was required when all our data
analyses occurred in batch mode on VAX (and later Unix) machines. The
fact that I need to articulate this principle at all arises because of
the rise of interactive data analysis and graphical user interfaces:
it is all too easy to use the mouse to load a data file into memory
and then to write a script to analyze this file without ever noting
the actual name or location of the data file.

\section{Version control prevents clobbering and reconciles history.}

Group work requires version control.\footnote{See X and Y in this
  issue for more discussion of what version control is.} Many people
are familiar with the ``track changes'' feature in modern WYSIWYG word
processors or the fact that Dropbox allows one to recover previous
versions of files. These are both kinds of version control. When
collaborating with yourself or others, it is useful to see what has
changed, to feel free to experiment and then to dump parts of the
experiment in favor of previous work while merging the successful
parts of the experiment into the main body of the paper, and to have
multiple ``releases'' of the same document (one to MPSA, one to APSR,
one to your parents) without spawning many possibly conflicting copies
of the same document, risking confusion and clobbering. Clobbering is
what happens when your future self or your current collaborator saves
an old version of a file over a new version --- erasing good work by
accident.

Of course if you rely on Dropbox or ``track changes'' for version
control, you must communicate with other folks in your group before
you edit existing files. Only one of you can edit and save a given
file at a time. This prevents your work (or your colleagues work) from
getting lost when you both try to save the same file on top of each
other. If you find that you need to work on the same files at the same
time, then you should work on establishing your own shared version
control system. Free options include launchpad, github, sourceforge
for open source projects (i.e. papers you are writing which you are
happy to share with others). Each of those services include paid
versions too. One may also use Dropbox as a kind of server for version
control: for example, one may copy files from the Dropbox directory
into a local working directory so as to avoid clobbering and then
working on merging changes by hand before copying over existing
files. (Notice that this is different from directly working on files
within your Dropbox managed directories.)

We use subversion with our own research group, and I use it for all of
my own projects (except this one, for which I am experimenting with
git). Subversion and bazaar and git are all great. They mainly differ
in the extent to which you need to run a server. Subversion requires a
server and we are lucky that the NCSA provides such a server for
us. However, the price of such hosting may not be that much using one
of the many webhosting services out there or perhaps may be available
for free at your university.\footnote{For example, if you already pay
  to have a website, you may already have the right to run a
  subversion server there.}

Fancy version control systems are not required, however, to get many
of the benefits of formal version control. I suspect that Google Docs
allows a kind of version tracking and collaboration as well.  An
excellent, simple, and robust version control system is to merely
rename your files with the date and time of saving them: thedoc.tex
becomes thedoc25-12-2011-23:50.tex.  Be sure to include year in the
file names --- remember, the life of an idea is measured in years. If
you are wise enough to have saved your documents as plain text then
you can easily compare documents using the many utilities available
for comparing text files. Adobe Acrobat allows one to compare
differences in pdf files. OpenOffice supports a ``Compare Documents''
option.

Also, if you use this method, spend a little extra time to
ensure that you do not clobber files when you make typos in the
file-names. And, you will find yourself spending extra time
reconciling changes made by different collaborators by hand that
modern version control systems take care of quickly and easily.

When you reach certain milestones you can rename the file accordingly:
thedocAPSA2009.tex --- for the one sent to discussants at APSA --- or
thedocAPSR2015.tex --- for the version eventually sent to the APSR six
years after you presented it at APSA. The formal version control
systems I mentioned above (and which are described in more depth in X
and Y) all allow this kind of thing and are much more elegant and
capable, but you can do it by hand too as long as you don't mind
taking up a lot of disk space and having  many ``thedoc...''
files around.

\paragraph{Step \thesection} Writing is rewriting. Thus, all writing involves
versions. When we collaborate with ourselves and others we want to
avoid clobbering and we want to enable graceful reconciliation of
rewriting. One can do these things with formal systems of software
(like subversion, git, etc...) or with formal systems of file naming,
file comparing, and communication or, even better, with both.

% This is so true that the
% source of the quote is unclear.\footnote{Could it be Paul Abbott?
%   \url{http://www.bbc.co.uk/writersroom/insight/paul_abbott.shtml}}
% Here is a quote with a known source: ``Writing and rewriting are a constant search for what it is one is
% saying'' (John Updike but ??Writing With Style: Conversations on the Art of Writing, 1975 
% by John Trimble ??)

\section{Minimize error by testing.}

Now, back to that confidence interval in the famous article of
2018. The statisticians at the UN worry about your use of the
bootstrap. The authors would like to evaluate their bootstrap
procedure --- and they did so originally in their memo comparing it to
a profile likelihood based approach. Although nice code exists for
bootstrapping linear models, no nice code exists to bootstrap the
bootstrap. Of course, the code required is not complex, but since they
are are writing custom code they worry about getting it right. Now, 9
years after the idea, they've had lots of time to appreciate problems
arising from bugs and errors in data analysis and code.

Now, if they had a moment to think in between teaching that new class,
reading books for the awards committee, reading application files for
the admissions committee, staying home with a sick child, and
undertaking the odd bit of your own current research, they might say to
themselves, ``Before I write new code, I should write a test of the
code. I should write a little bit of code that let's me know that my
double-bootstrap procedure actually does what it is supposed to do.''

Of course, this idea, like most others, is not new. When large groups
of programmers write code for multi-million dollar programs the
question about avoiding error looms large. The idea of
\href{http://en.wikipedia.org/wiki/Test-driven_development}{test
  driven development} and the idea that one ought to create tests of
\href{http://en.wikipedia.org/wiki/Unit_testing}{small parts of one's
  code} arose to address such concerns. For the social scientist
collaborating with her future self and/or a small group of
collaborators here is an example of this idea in a very simple form. I
want to write a function to multiply a number by 2. If my function
works, when I give it the number 4, I should see it return the number
8 and when I give it -4, I should get -8.

<<unit.test>>=
##  The test function:
test.times.2.fn<-function(){
  ##  This function tests times.2.fn
  if (times.2.fn(thenumber=4) == 8 &
     times.2.fn(thenumber=-4) == -8) { 
    print("It works!")
  } else { print("It does not work!")
         }
}

##  The function:
times.2.fn<-function(thenumber){
  ##  This function multiplies a scalar number by 2
  ##  thenumber is a scalar number
  thenumber+2
}

##Use the test function
test.times.2.fn()
@

Ack! I mistyped ``+'' for ``*''. Good thing I wrote the test!

\paragraph{Step \thesection} You cannot forsee all of the ways that your code
could be used, but you can at least make sure it does what it is
supposed to do in your particular case.

\section{Copy and improve on others' examples.}

Lots of people are thinking about ``reproducible research'' and
``literate programming'' these days. Google those terms. Of course the
devil is the details: Here I list a few of my own attempts at enabling
reproducible research. You'll find many other inspiring examples on
the web. Luckily, the open source ethos aligns nicely with academic
incentives, so we are beginning to find more and more people offering
their files online for copying and improvement. By the way, if you do
copy and improve, it is polite to alert the person from whom you made
the copy about your work.

I have experimented with three systems so far: (1) for one paper we
simply included a Sweave document and data files into a compressed
archive \citep{bowers2005dataverse}; (2) for another more computing
intensive paper we assembled a set of files which enabled reproduction
of our results using the ``make'' system \citep{bowers2008dataverse};
and (3) recently I have tried the ``compendium'' approach
\citep{gentleman2005reproducible,gentleman2007statistical} which
embeds an academic paper with the R package system
\cite{bowers2011dataverse}. The benefit of this last system is that
one is not required to have access to a command line for \Verb+make+:
the compendium is downloadable from within R using
\Verb+install.packages()+ and it viewable using the \Verb+vignette()+
function. The idea that one ought to be able to install and run and
use an academic paper just as one installs and uses statistical
software packages is very attractive and I anticipate that it will
become ever easier to turn papers into R packages as creative and
energetic folks turn their attention to the question of reproducible
research.

\section{Remember that research ought to be credibile communication.}
\begin{quote}
  [I]f the empirical basis for an article or book cannot be
  reproduced, of what use to the discipline are its
  conclusions?  What purpose does an article like this serve?
  \cite[445]{king1995replication}
\end{quote}

We all always collaborate. Many of us collaborate with groups of
people at one moment in time as we race against a deadline. All of us
collaborate with ourselves over time.\footnote{What is a reasonable
  time-span for which to plan for self-collaboration on a single idea?
  Ask your advisers how long it took them from idea to dissertation to
  publication.}  The time-frames over which collaboration are required
--- whether among a group of people working together or within a
single scholar's productive life or probably both --- are much longer
than any given version of any given software will easily exist. Plain
text is the exception. Thus, even as we extol version control systems,
one must have a way to ensure future access to them in a form that
will still be around when sentient cockroaches finally join political
science departments (by then dominated by cetaceans after humans are
mostly uploads).\footnote{The arrival of the six-legged social
  scientists revives Emacs and finally makes Ctrl-a Ctrl-x Esc-x
  Ctrl-c a \href{http://kieran.healy.usesthis.com/}{reasonable key
    combination}.}

But what if the UN never hears of your work, or, by some cruel fate, that your
article does not spawn debate? Why then would you spend time to
communicate with your future self and others? My own answer to this
question is that I want my work to be credible and useful to myself
and other scholars even if each article does not immediately change
the world.  What I report in my data analyses should have two main
characteristics: (1) the findings of the work should not be a matter
of opinion; and, (2) other people should be able to reproduce the findings. That is,
the work represents a shared
experience --- and an experience shared without respect to the
identities of others (although requiring some common technical
training and research resources).
% Such work should make us change how
% we act --- or at least, it ought to stand on stronger epistemological
% ground than other claims about experiences which ought to be shared or
% shareable.

Assume we want others to believe us when we say something. More
narrowly, assume we want other people to believe us when we say
something about data: ``data'' here can be words, numbers, musical
notes, images, ideas, etc \ldots The point is that we are making some
claims about patterns in some collection of stuff. Now, it might be
easy to convince others that ``this collection of stuff is different
from this collection of stuff'' if those people were looking over our
shoulders the whole time that we made decisions about collecting the
stuff and broke it up into understandable parts and reorganized and
summarized it. Unfortunately, we can't assume that people are willing
to shadow a researcher throughout her career. Rather, we do our work
alone or in small groups and want to convince other distant and future
people about our analyses.

Now, say your collections of stuff are large or complex and your
chosen tools of analyses are computer programs. How can we convince
people that what we did with some data with some program is credible:
not a matter of whim or opinion, and reproducible by others who didn't
shadow us as we wrote our papers? This essay has suggested a few
concrete ways to enhance the believeability of such scholarly work. In
addition, these actions (as summarized in the section headings of this
essay) make collaboration within research groups more effective.
Believeability comes in part from reproducibility and research groups
often need to be able to reproduce in part or in whole what different
people in the group have done.

In the end, following these practices and those recommended by X and Y
in this issue allows your computerized analyses of your collections
of stuff to be credible.  Finally, if the UN quibbles with your
analyses, your future self can shoot the archive required to reproduce
your work (in still intelligible plain text, analyzed using commented
code so that folks can translate to whatever system succeeds R, or
since you used R, you can include a copy of R and all of the R
packages you used in your final analyses in 2018 in the archive
itself.) You can say, ``Here is everything you need to reproduce my
work." To be extra helpful you can add ``Read the README file for
futher instructions." And then you can get on with your life: maybe
the next great idea will occur when your 4-year-old asks a wacky
question after stripping and painting her overly cooperative
1-year-old brother purple, or teaching a class, or in a coffee shop,
or on a quiet walk.

\bibliographystyle{apsr}
\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}

\end{document}
<<>>=
options(prompt="> ",continue="+ ")
@