### 2016-12-07 workflow.Rmd
made some edits

---
title: How to improve your relationship with your future self
author:
- name: Jake Bowers
  affiliation: <jwbowers@illinois.edu>,  University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team.
- name: Maarten Voors
  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group.  \emph{Acknowledgements:} Many thanks to the EGAP Learning Days 2016 participants in Santiago Chile, to the BITSS team, the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016. Maarten gratefully acknowledges financial support from N.W.O. grant 451-14-001. A previous version of this paper benefited from comments and discussions with Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn, Cara Wong, Mika LaVaque-Manty and Ben Hansen. The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>. This paper extends a previous version by Jake @bowers2011tpm. While some text remains the same we have expanded and updated the essay to reflect current developments.
bibliography: references.bib
csl: american-political-science-review.csl
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
    pdf_document:
      keep_tex: true
      number_sections: true
      fig_width: 7
      fig_height: 7
      fig_caption: true
      template: bowersarticletemplate.latex
      pandoc_args: [
      "-M", "secPrefix=\\S",
      "--filter", "/usr/local/bin/pandoc-crossref"
      ]
      md_extensions: +autolink_bare_uris+ascii_identifiers+tex_math_single_backslash
...

<!-- An example comment -->
<!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->

```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

## To make the html file do
## render("exploration4.Rmd",output_format=html_document(fig_retina=FALSE))
## To make the pdf file do
## render("exploration4.Rmd",output_format=pdf_document())
## to debug, try render("thefile.Rmd", clean=FALSE) so that you can see the intermediate files

require(knitr)
opts_chunk$set(
  tidy=TRUE,     # display code as typed
  size="footnotesize",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=100,
  out.width='.9\\textwidth',
  message=FALSE,
  comment=NA)
```

```{r initialize,echo=FALSE}
##First, just setup the R environment for today:
if(!file.exists('figs')) dir.create('figs')

options(SweaveHooks=list(fig=function(){
			   par(mar=c(3.5, 3, 1.1, 0),
			       pty="s",
			       mgp=c(1.5,0.5,0),
			       oma=c(0,0,0,0))},
			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
			 options(prompt=" ")
			 }),
	digits=4,
	scipen=8,
	width=100
	)

#options(error=function(){options(prompt="> ",continue="+ ");NULL})
```

> If you tell the truth, you don't have to remember anything. [@twain1975mark]

Memory is tricky. Learning requires effort. When we do not practice and repeat something that we want to remember, most people forget quickly, are overconfident in their abilities to recall future information [@koriat2005illusions], and may even recall events that never happened.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>. On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] Moreover, most of us live busy lives. We write text and code behind our new standing desk at home while the kids run around, cry about an elder sister taking away a favorite new yellow school bus toy, drinks spill on the floor, the laundry needs doing, the kettle boils \ldots Our lives and minds are full and we need to efficiently move from task to task. If we cannot count on memory, then how can we do science?

How long does it take from planning a study to publication? How much time passes to the first reproduction of a study? Is three years too short? Is ten years too long? We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a data driven study faster than about three years. We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication. Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, many years past the initial publication of the article.^[The process of reproducing past findings can occur when one researcher wants to build on the work of another. It can also occur within the context of classes --- some professors assign reproduction tasks to students to aid learning about data analysis and statistics. In addition to those models, recently reproduction of research has been organized to enhance the quality of public policy in the field of economic development by the 3IE Replication Program (@brown2014quality) and to assess the quality of scientific research at the level of sub-disciplines in social psychology (@open2015estimating) and within experimental economics (@camerer2016evaluating). In another study 29 research teams recently collaborated on a project focusing on applied statistics to see if the same answers would emerge from re-analyses of the same data set (@silberzahn2015crowdsourced).]

If we cannot count on our memories about why we made such and such a design or analysis decision, then what should we do? How can we minimize regret with our past decisions? How can we improve our relationship with our future self?  This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make reproducible data analysis easy and quick. Specifically, this piece aims to amplify some of what we already ought to know and do, and highlight some current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists. Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT initiative <http://www.dartstatement.org/> @lupia2014openness. These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations, <http://osf.io> and <http://www.bitss.org/>.] We aim to provide practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on bad past decisions that do not have to do with data analysis and the production of scholarly papers.

```{r read.data,echo=FALSE,results='hide', cache=TRUE}
## This data frame will be used throughout the analyses.
## Putting it up top, so we can use it in the comments section.
library(readstata13)
dem.nations.df<-read.dta13("https://www.dropbox.com/s/huhqag5rudwtbno/Democracy%20Cross-National%20Data%20V4.1%2009092015.dta?dl=1&pv=1&raw=1",convert.factors=FALSE)

row.names(dem.nations.df)<-dem.nations.df$Nation ##make the row names more helpful
good.df<-na.omit(dem.nations.df[,c("Gini2004","WVS5_protest","Nation","meanpr")])
good.df$gini04<-good.df$Gini2004/100
good.df$protest05<-good.df$WVS5_protest
```

We organize the paper around a series of homilies that lead to certain concrete actions.

# Data analysis is computer programming. {#sec:data-analys-comp}

All your results (numbers, comparisons, tables, plots, figures) should be produced from
code, not from a series of mouse clicks or copying and pasting.^[To our future
selves: both of us are using a computer control device called a track-pad, and
haven't used an older device called a
[mouse](https://en.wikipedia.org/wiki/Computer_mouse) in some years. We no
longer hear a clicking sound when we operate the control device either. We are
not sure why we talked about mouse clicks just now.] Imagine you wanted to re-create
a figure and include a new variable, you should be able to do so with just a few edits to the code rather than knowledge
of how you used a pointing device in your graphical user interface all those
years ago.

Lets look at an example. Using R [@R:2011], a open-source statistical software program, you
might specify that a file, called `fig1.pdf` is produced by the following set
of commands in a file called `makefig1.R`. Lets look at some annotated R code:

```{r eval=FALSE,tidy=FALSE}
## This file produces a plot relating the explanatory variable to the outcome.
thedata <- read.csv("Data/thedata-15-03-2011.csv") ## Read the data
pdf("fig1.pdf") ## begin writing to the pdf file
please-plot(outcome by explanatory using thedata. red lines please.)
please-add-a-line(using model1)
## Note to self: a quadratic term does not add to the substance
## model2 <- please-fit(outcome by explanatory+explanatory^2 using thedata
## summary(abs(fitted(model1)-fitted(model2)))
dev.off() ## stop writing to the pdf file
```

Now, in the future if you wonder how 'that plot on page 10' was created, you
will know: (1) 'that plot' is from a file called `fig1.pdf`and (2)
`fig1.pdf`was created in `makefig1.R`. Any changes to the file in the future (provided R still
exists), will just require some quick edits of commands already
written.^[We use data from @norris2015data throughout this paper.]  Even if in the future R ceases to exist, you (or someone else) will at
least be able to read the plain text commands and use them to write code in a
new favorite statistical computing language: R scripts are written in [plain
text](http://en.wikipedia.org/wiki/Plain_text), and plain text is a format that
will be around as long as computer programmers write computer programs.

Moreover, coding saves time. Often, a lot of time, especially if projects grow
in the number of files, partners, or complexity of analysis. Manually importing
one data file into R, may be effortless. Importing 100 files is another issue,
and the time costs of each manual action add up quickly (and the probability
for error increases too).

Also, realize that *file names send messages* to coauthors and to your future self. This means
that if you name your files with evocative and descriptive names, your
collaborators are less likely to call you at midnight asking for help and you
will remove some regret from your future self and protect your friendships and
working relationships. For example, if you are studying inequality and protest,
you might try naming a file something like `inequality-and-protest-figures.R`
instead of `temp9.R` or `supercalifragilisticexpialidocious.R`.  By the way,
the extension `.R` tells us and the operating system that the file contains R
commands. This part of the filename enables us to quickly search our antique
hard drives for files containing R scripts.^[We think that some method of
tagging files by the purpose of the file will continue help analysts find and organize their files for long after the idea of a computer mouse ceases to make sense.]

Coding helps us to avoid making mistakes. For example, in our example above we may
be interested in how many people protest. We may use a data file containing all
protests for several years. People often create a tabular display of this data
by copying the results manually to the working paper document. In copying we
can make mistakes. So a better approach is to automatically create the table (in the format you like, with
horizontal lines, 3 decimals, a title, etc) and output this in whatever file
type you need (.tex. .pdf, .rtf, etc). Now when we obtain new data a new table can be created
quickly, so we make less mistakes and save time!

For example, the following table was created entirely with code using R and the `xtable` package @dahl2016aa.\footnote{@beck2010reg inspired this particular
presentation of a linear model.}

```{r loadlibraries, echo=FALSE}
makebecktable<-function(lmobj,vars,thecaption,thelabel,filename){
	require(xtable)
	##  The next lines extract relevant and useful output from the
	##  regression object, following Neal Beck's TPM article 2010.
	themat<-cbind(summary(lmobj)$coef[,c("Estimate","Std. Error")],confint(lmobj))
	n<-length(lmobj$fitted.values)
	r2<-signif(summary(lmobj)$r.squared,2)
	est.sigma<-signif(summary(lmobj)$sigma,2)
	colnames(themat)[3:4]<-c("~{}","~{}")
	row.names(themat)<-vars
	thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
			caption=thecaption,
			label=thelabel)
	print(thextab,
	      file=filename,
	      include.colnames=FALSE,
	      hline.after=NULL,
	      table.placement="H",
	      comment=FALSE,
	      sanitize.text.function=function(x){x},
	      add.to.row=list(
			      pos=list(0,3),
			      command=c(
					"\\toprule \n &
					\\multicolumn{1}{c}{Coef} &
					\\multicolumn{1}{c}{Std. Err.} &
					\\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
				paste("& \\multicolumn{4}{c}{n: ",n,",",
				      "   resid.sd: ",signif(est.sigma),
				      ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
				)
			      ))

}

```

```{r maketable1, echo=TRUE, tidy=FALSE}
lm1<-lm(protest05~gini04+meanpr,data=good.df)  #  Run the regression
makebecktable(lmobj=lm1, # make the table file
	      vars=c("Intercept","Income Inequality (lower=more equal)",
		     "Mean Political Rights (lower=more rights)"),
	      thecaption="People living in countries with unequal
	      income distributions report more protest activity to World Values
	      Survey interviewers than people living in countries with relatively
	      more equal income distributions, adjusting for average political
	      rights as measured by Freedom House 1980--2010.",
	      thelabel="tab:protest",
	      filename="protesttable.tex")
```

```
\input{protesttable.tex} ## read in the table file
```

\input{protesttable.tex}

**Step 1** Code everything that can be coded. If we know the provenance of results, future or current collaborators make less mistakes and can quickly and easily reproduce (and thus change and improve) upon the work.

# No data analyst is an island for long.

Data analysis involves a long series of decisions. Each decision requires a
justification.  Some decisions will be too small and technical for inclusion in
the published article itself. Still, these need to be documented in the code
itself [@nagler1995coding]. Paragraphs and citations in the publication will
justify the most important decisions but the code itself documents the smaller
but still consequential decisions. So, one must code to communicate with
yourself and others. There are two main ways to avoid forgetting the reasons
you did something with data: comment your code and tightly link your code with
your writing -- making your code literate.

## Code is communication: Comment code

Comments, unexecuted text inside of a script, are a message to collaborators
(including your future self) and other consumers of your work. In the above
code chunk, we used comments to explain the lines to readers unfamiliar with R
and to remember that we had tried a different specification but decided not to
use it because adding the squared term did not really change the substantive
story arising from the model.^[R considers text marked with `#` as a comment.
For Stata simply add a `*` before the text.] Messages left for your future self
(or near-future others) help retrace and justify your decisions as the work
moves from seminar paper to conference paper to poster back to paper to
dissertation and onwards maybe even to publication.

Notice one other benefit of coding for an audience: we learn by teaching. By
assuming that others will look at your code, you will be more likely to write
clearer code, or perhaps even to think more deeply about what you are doing as
you do it since you are explaining even as you write.

Comment liberally. Comments are discarded when R or STATA runs analysis, so
only those who dig into the source code of your work will see them.

## Code to communicate: Literate programming.

> Let us change our traditional attitude to the construction of
  programs: Instead of imagining that our main task is to instruct a
  computer what to do, let us concentrate rather on explaining to
  human beings what we want a computer to do.
  [@knuth1984literate, p.97]

Imagine you discover something new (or confirm something old). You
produce a nice little report on your work for use in discussions of
your working group or as a memo for a web or reviewer appendix. The
report itself is a pdf or html file or some other format which displays page
images to ease reading rather than to encourage reanalysis and
rewriting. Eventually pieces of that report (tables, graphs,
paragraphs) ought to show up in, or at least inform, the publishable
paper. Re-creating those analyses by pointing, clicking, copying, or
pasting would invite typing error and waste time. Re-creating your
arguments justifying your analysis decisions would also waste
time.

More importantly, we and others want to know why we did what we did. Such
explanations may not be very clear if we have some pages of printed code in one
hand and a manuscript in the other. Keep in mind the distinction between the
'source code' of a document (i.e. what computation was required to produce it)
and the visible, type-set page image. Page images are great for reading, but
not great for reproducing or collaborating. The source code of any document
exchanged by the group must be available and executable.^[As of the 2016
version of this paper, this idea is now widespread and made much easier than
before via online services for code sharing and collaboration such as
[GitHub](http://github.com), [Open Science Framework](https://osf.io/) and
[BitBucket](http://bitbucket.org). As more data analysis moves online, it will
become easier for cross-platform and geographically distant collaboration to
occur, see for only one set of examples, [Docker](http://docker.com) or other
services that make what is currently called 'cloud-computing' easier and more
accessible.]

How might one avoid these problems?  [Literate
programming](http://en.wikipedia.org/wiki/Literate_programming) is the practice
of weaving code into a document --- paragraphs, equations, and diagrams can
explain the code, and the code can produce numbers, figures, and tables (and
diagrams and even equations and paragraphs). Literate programming is not merely
fancy commenting but is about enabling the practice of programming itself to
facilitate easy reproduction and communication.

For example, in [@sec:data-analys-comp], we suggested that we knew where 'that
plot on page 10' comes from by making sure we had a `fig1.pdf` file produced
from a clearly commented plain text file called something like `makefig1.R`.
An even easier solution would be to directly include a chunk of code to produce
the figure inside of the paper itself. This paper, for example, was written in
plain text using markdown markup with R code chunks to make things like Figure
\ref{fig:giniprot}. This combination of markdown and R is called [R
Markdown](http://rmarkdown.rstudio.org). Here, is an example of embedding a
plot inside of a document. In a document sent to a journal one would tend to hide the code chunks and thus one would only see the plot.

```{r fig1plotcode,eval=FALSE,echo=TRUE}
##  Make a scatterplot of Protest by Inequality
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
with(good.df,plot(gini04,protest05,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
                  cex=.8))
##  Label a few interesting points
with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
```

```{r giniprot,message=FALSE,echo=FALSE,fig=TRUE,include=TRUE,width=4,height=4,fig.cap="Average number of protest activities by income inequality across countries in 2004--2005.",out.width=".5\\textwidth",fig.env="figure",fig.lp="fig:",fig.pos="H",fig.align="center"}
##  Make a scatterplot of Protest by Inequality
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
with(good.df,plot(gini04,protest05,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
                  cex=.8))
##  Label a few interesting points
with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
```

Markdown (and LaTeX and HTML) all have ways to cross-reference within a
document. For example, by using `\label{fig:giniprot}` in \LaTeX or
`{#fig:giniprot}` in the pandoc version of markdown, we do not need to keep
track of the figure number, nor do extra work when we reorganize the document
in response to reviewer suggestions. Nor do we need a separate `makefig1.R`file
or `fig1.pdf` file.  Tables and other numerical results are also possible to
generate within the source code of a scholarly paper. For example, if we had
ommitted the filename in `makebecktable` above, we would have seen a LaTeX
table be generated within this document itself. For Stata users, there now is
[Markdoc](https://github.com/haghish/MarkDoc) which is very similar to R
markdown.

The R project has a task view devoted to [reproducible
research](http://cran.r-project.org/web/views/ReproducibleResearch.html)
listing many of the different approaches to literate programming for R.  If
your workflow does not involve R, you can still implement some of the
principles here. Imagine creating a style in Microsoft Word called 'code 'which
hides your code when you print your document, but which allows you to at least
run each code chunk piece by piece (or perhaps there are ways to extract all
text of style code from a Microsoft Word document). Or imagine just using some
other kind of indication linking paragraphs to specific places in code files.
There are many ways that creative people can program in a literate way.

Literate programming need not go against the principle of modular data analysis
[@nagler1995coding]. Jake routinely uses several different files that fulfill
different functions, some of them create \LaTeX code that he can `\input` into
his `main.tex` file, others setup the data, run simulations, or allow him to
record his journeys down blind alleys. Of course, when we have flying cars
running on autopilot, perhaps something other than combining R and Markdown or
LaTeX will make our lives even easier. Then we'll change.

**Step 2** We analyze data in order to explain something about the world to
other scholars and policy makers. If we focus on explaining how we got our
computers to do data analysis, we will do a better job with the data analysis
itself: we will learn as we focus on teaching others about why we did what we
did, and we will avoid errors and save time as we ensure that others (including
our future selves) can retrace our steps. A document than can be 'run' to
reproduce all of the analyses also instills confidence in readers and can more
effectively spur discussion and learning and cumulation of research.

# Meaningful code requires data.

All files containing commands operating on data must refer to a data file. A
reference to a data file is a line of code the analysis program will use to
operate on ('load'/ 'open '/ 'get '/ 'use') the data file. One should not have
to edit this line on different computers or platforms in order to execute this
command.  Using R, for example, all analysis files should have
`load("thedata.rda")`or `read.csv("thedata.csv")`or some equivalent line in
them, and `thedata.csv`should be stored in some easy to find place (like in the
same directory as the file or perhaps in `"Data/thedata.rda"`). Of course, its
even better to include a comment pointing to the data file in addition to the
line loading the file itself.

Where should one store data files? An obvious solution is to
make sure that the data file used by a command file is in the same
directory as the command file. More elegant solutions require all
co-authors to have the same directory structure so that
`load("Data/thedata.rda")` means the same thing on all computers
used to work on the project. This kind of solution is one of the
things that Dropbox and more formal version control systems do well (as discussed
a bit more below in [@sec:vers-contr-prev]).

The principle of modularity suggests that you separate data cleaning,
processing, recoding, and merging from analysis in different files
[@nagler1995coding]. So, perhaps your analysis oriented files will
`load("cleandata.rda")`and a comment in the code will alert the future you
(among others) that `cleandata.rda`was created from `create-cleandata.R` which
in turn begins with
`read.csv(\url("http://www.greatfreedata.gov/dirtydata.csv"))`. Such a data
processing file will typically end with something like
`save("cleandata.rda")`so that we are doubly certain about the provenance of
the data.^[Of course, if you need math or paragraphs to explain what is
happening in these files, you might prefer to make them into R+Markdown or
R+LaTeX files, for which the conventional extension is .Rmd or .Rnw
respectively. So you'd have `create-cleandata.Rnw` written as a mixture of
LaTeX and R which might explain and explore the different coding decisions you
made, perhaps involving some diagnostic plots.]

Now, if in the future we wonder where `cleandata.rda`came from, we
might search for occurrences of `cleandata` in the files on our
system. However, if such searching among files is a burden, an even
nicer solution is to maintain a file for each project called
MANIFEST.txt or INDEX.txt or README.txt which lists the
data and command files with brief descriptions of their functions and
relations.

Best practice is to create a file system where you separate folders by function and separate input from output files. For example, below we show a folder structure for a paper where we look at inequality. The paper is written in .tex, data analysis in Stata (that Maarten prefers, so we see .do and .dta file extensions) in R (the program of choice for Jake, see the .R extensions). Maarten likes to make sure that the directory structure in his projects is the same across projects (so, for example, the numbers on the directory names allow for easy default sorting). The file naming with full dates also allows for easy sorting:

```{bash ,eval=FALSE, tidy=FALSE}
00_archive/
01_paper/
  20160622_inequality.tex
02_data/
  00_archive/
  01_rawdata/
  02_cleandata/
  20160622_analysis.dta
  20160622_analysis.rda
03_analysis/
  00_archive/
  01_temp/
  02_output/
    01_tables/
    02_figures/
  20160622_analysis.do
  20160622_analysis.R
  20160622_prep_data.do
  20160622_prep_data.R
 ```

There are a couple of things to note: See that files start with a number to give them an ordering. Note also we create an 00_archive folder, here you can store older files without having to delete them permanently and can reduce the number of files in your main folders. Also there is a clear separation between raw data (data that came straight from the field, downloaded data, someone else's replication files, etc) and clean data (data ready to use). Note that in the analysis folder we created two files, one for cleaning and merging data called `20160622_prep_data` and another for running the analysis called `20160622_analysis`. See also the output folders, with subfolders for graphs and tables. These are place holder for R or STATA to output the results from the analysis and keeps things organized.

Jake tends to be less organized than Maarten and relies on README.md files to tell his future self what is going on in a given directory, a [version control system](http://github.com) to keep track of versions of files, and a [Makefile](http://kbroman.org/minimal_make/) to keep track of relationships between files.

```{bash, tidy=FALSE, eval=FALSE}
manuscript\
figures\
analysis\
  README.md
data\
build\
libraries\
README.md
Makefile
```

**Step 3** We should know where the data came from and what operations were performed on which set of data.

# Version control prevents clobbering, reconciles history, and helps organize work. {#sec:vers-contr-prev}

Group work requires version control. Many people are familiar with the 'track
changes' feature in modern WYSIWYG (what you see is what you get) word
processors or the fact that Dropbox allows one to recover previous versions of
files. These are both kinds of version control.  More generally, when we
collaborate, we'd like to do a variety of actions with our shared files.
Collaboration on data analytic projects is more productive and better when (1)
it is easy to see what has changed between versions of files; (2) members of
the team feel free to experiment and then to dump parts of the experimentation
in favor of previous work while merging the successful parts into the main body
of the paper; (3) the team can produce 'releases' of the same document
(one to MPSA, one to APSR, one to their parents) without spawning many possibly
conflicting copies of the same document; (4) people can work on the same files
at the same time without conflicting with one another, and can reconcile their
changes without too much confusion and clobbering. Clobbering is what happens
when your future self or your current collaborator saves an old version of a
file over a new version, erasing good work by accident (or creating a conflicted copy resulting in an arduous back and forth between versions).

Using Dropbox, 'track changes' in Word, or Google Docs is one way to manage your collaborative efforts. They are popular tools, but
have some downsides as they require you to communicate with other folks
in your group before you can edit existing files. On Dropbox, you cannot
edit and save a given file at the same time. Tracking changes with Word has similar limitations (merging the
changes between two documents is not simple or builtin). Google Docs allows for
two people to edit the same document at the same time: but you must be online
with a fairly fast internet connection.^[Note to future selves: change this
section every five years or so. Probably stop saying 'internet' in 2026.] Communication
and trading off editing a file prevents your work (or your colleagues' work)
from getting lost when you both try to save the same file on top of each other. One may also use Dropbox as a kind of server for version
control. See example above. This system is fine untill your collaborations or project grows large
(in terms of MBs) and you may run out of storage space on your computer or you may need to upgrade your Dropbox's storage space). You will however need to agree on an iron rule for file and directory naming with your collaborators to ensure that the files and directories are always named the same across machines and versions (see above).

An excellent, simple, and robust version control system is to rename your files
with the date and time of saving them: yyyymmdd_project.docx (for changes
within the same day, just before a deadline for example, you may even add a
time, so it becomes 20160623_4PM_inequalitypaper.docx). This communicates very
clear what version you are working. Remember the days when you received a file
called inequalityMV4_APSRsubmit_reallyfinal2.tex? That should quickly become
something of the past! We find ourselves preaching the gospel to our
collaborators frequently and will keep renaming files until a collaborator has
been successfully pacified. Be sure to include year in the file names ---
remember, the life of an idea is measured in years. If you are wise enough to
have saved your documents as plain text then you can easily compare documents
using the many utilities available for comparing text files.^[Adobe Acrobat
allows one to compare differences in pdf files. OpenOffice supports a 'Compare
Documents' option. Word now does the same. And Google Docs will report on the
version history of a document.] When you reach certain milestones you can
rename the file accordingly: thedocAPSA2009.tex --- for the one sent to
discussants at APSA --- or thedocAPSR2015.tex --- for the version eventually
sent to the APSR six years after you presented it at APSA. The formal version
control systems mentioned above all allow this kind of thing and are much more
elegant and capable, but you can do it by hand too as long as you don't mind
taking up a lot of disk space and having many 'thedoc... ' files around.
Indeed, the number of files can add up quickly (especially around submission
time!) and you may want to create an "0_Archive" folder to store older
versions. If you do version control by hand, spend a little extra time to
ensure that you do not clobber files when you make mistakes typing in the
file-names. And, if you find yourself spending extra time reconciling changes
made by different collaborators by hand, remember this is a task that modern
version control systems take care of quickly and easily.

However, best practice we feel is to use formal version control approaches to organizing work [version
control](https://en.wikipedia.org/wiki/Version_control) strategy. As of 2016,
the standard for managing large collaborations with many files is the
[Git](https://en.wikipedia.org/wiki/Git_(software)) system. User friendly free
front-ends for Git currently make the process of learning and using Git very
convenient and integrated into the kind of workflow that your future self will
be proud of: [Open Science Framework](http://osf.io)
[bitbucket](http://bitbucket.com), [github](http://github.com). Jake uses
[github](http://github.com) for all of his collaborations with others as well
as with his future self: his collaborators appreciate some of the nice extra
features of GitHub, such as the ability to keep a [shared task
list](https://guides.github.com/features/issues/). Learning about version
control systems takes a bit of time. We suggest, however, it is well worth the
time investment as it will save (lots!) of time later on.

**Step 5** Writing is rewriting. Thus, all writing involves versions. When we
collaborate with ourselves and others we want to avoid clobbering and we want
to enable graceful reconciliation of rewriting. One can do these things with
formal systems of software (like git) or with formal systems of file naming,
file comparing, and communication or, even better, with both. In either case,
plain text files will make such tasks easier, will take up less disk space, and
be easier to read for the future you.

# Minimize error by testing.

Anyone writing custom code should worry about getting it right. The more code
one writes, the more time one has to appreciate problems arising from bugs,
errors, and typos in data analysis and code. Since mistakes will always be
made, we should just recognize we need more standardization of practice to
minimize the mistakes and help others find them. One thing we can do to
minimize and catch the inevitable mistakes is to include testing in our coding.

Of course, this idea, is not new. The desire to avoid error
looms large when large groups of programmers write code for multi-million
dollar programs. The idea of [test driven
development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea
that one ought to create tests of [small parts of one's
code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such
concerns.^[There are now R packages to help R package writers do this, see for
example <https://github.com/hadley/testthat> and the article on it
<https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf>.]

For the social scientist collaborating with her future self and/or a small
group of collaborators, here is an example of this idea in a very simple form:
Say you want to write a function to multiply a number by 2. If the function
works, when you give it the number 4, you should see it return the number 8 and
when you give it -4, you should get -8.

```{r unit.test, tidy=FALSE}
##  The test function:
test.times.2.fn <- function(){
	##  This function tests times.2.fn
	if (times.2.fn(thenumber=4) ==  8 &
	    times.2.fn(thenumber=-4) == -8) {
		print("It works!")
		} else { print("It does not work!")
	}
}

##  The actual function is:
times.2.fn <- function(thenumber){
	##  This function multiplies a scalar number by 2
	##  thenumber is a scalar number
	thenumber+2
}
```
Here we use the test function to make sure it works.

```{r results='asis'}
test.times.2.fn()
```

Ack! we mistyped '+' for '*'. Good thing we wrote the
test!^[A more common example of this kind of testing occurs
everyday when we recode variables into new forms but look at a
crosstab of the old vs. new variable before proceeding.]

That approach works well when one is paying close attention to messages within
the code. Another approach that we like better is to make the code stop with an
error whenever it doesn't pass a test. In R we use the `stopifnot` function for
this. For example, say we wanted to re-run our analyses excluding the countries
with extreme values of protest. The following code makes a new dataset
excluding places where more than 90% of the respondents to the World Values
survey reported some protest activity. If all of the values of `protest05` are
not less than 90, the code will stop with an error.

```{r}
smalldat <- subset(good.df,subset=protest05<90)
stopifnot(all(smalldat$protest05<90))
```

**Step 6** No one can forsee all of the ways that a
computer program can fail. One can, however, at least make sure that
it succeeds in doing the task motivating the writing of the code in
the first place.

# Create a reproducible workflow

Over the past decades more and more attention is being paid to 'reproducible research', creating a
'reproducible workflow' and 'literate programming' these days. As always, the devil is in the details: Here we list a few of our own
attempts at enabling reproducible research. You'll find many other inspiring
examples on the web. Luckily, the open source ethos aligns nicely with academic
incentives, so we are beginning to find more and more people offering their
files online for copying and improvement. By the way, if you do copy and
improve, it is polite to alert the person from whom you made the copy about
your work and to credit them in some way.

We have experimented with several systems so far:

(1) We wrote this paper in the [R markdown](http://rmarkdown.rstudio.com/)
literate programming format using two different text editors (Jake used vim)
and (Maarten used [atom](https://atom.io/)). We organized our files and
collaboration on GitHub. The source code for this document can be accessed,
downloaded and modified from <https://github.com/jwbowers/workflow>. We also
used GitHub Issues to send notes to each other and maintain a task list. The
nice thing about GitHub is that is enables you to make "releases" of a project
which enable you to smooth the reproduction of all of the products of your
research (simulations, data analyses, tables, figures, etc..).  Similar
features are offered by the [Open Science Framework (OSF)](https://osf.io/).
Maarten recently participated in a workshop on research transparency organized
by the [Berkeley Institute for Transparency in the Social
Sciences](www.bitss.org) in which all [course material](https://osf.io/qh2nr/)
was accessible through OSF.  OSF integrates with Github nicely and a key
benefit is that both systems have been created to remain for a long time
(unlike perhaps your university drive).  (2) For one paper Jake simply mixed R
and \LaTeX code in one document (called the "Sweave" format of literate
programming) and then added that document and data into into a compressed
archive [@bowers2005dataverse]; (3) For another more computing intensive paper
Jake assembled a set of files that enabled reproduction of results using the
`make` system [@bowers2008dataverse]; (4) Jake also tried the "compendium"
approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
academic paper within the R package system [@bowers2011dataverse]; The benefit
of the compendium approach is that one is not required to have access to a
command line for `make`: the compendium is downloadable from within R using
`install.packages()` and is viewable using the `vignette()` function in any
operating system than runs R.^[Notice that Jake's reproduction archives and/or
instructions for using them are hosted on the
[Dataverse](http://thedata.org/book/learn-about-project), which is another
system designed to enhance academic collaboration across time and space.]  The
idea that one ought to be able to install and run and use an academic paper
just as one installs and uses statistical software packages is very attractive,
and we anticipate that it will become ever easier to turn papers into R
packages as creative and energetic folks turn their attention to the question
of reproducible research.^[See for example <http://r-pkgs.had.co.nz/>] Finally,
(5) Maarten has been using Dropbox for most of his collaborative projects. See
comments above. As a tool for making files publicly available Dropbox is less
useful. You can of course make replication files available through creating an
public folder, but you will still need to refer to this folder on a particular
project website (an example is [Maarten's project on conflict and football](
https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl=0)).

There are some noteworthy new developments also. See for example
a new web app, Jupyter Notebook (http://jupyter.org/), that enables
you to make and share documents that include text, code, equations and
visualizations in one [literate programming
environment](http://blog.jupyter.org/2015/05/07/rendering-notebooks-on-github/).
For researchers working across different platforms, there is
[Docker](https://www.docker.com/) which promises to enable all of the
collaborators on one project to use the same computing environment even if they
are using different laptops running different operating systems.

There are also are great notes programs that help you communicate with collaborators. Remember, while email is fantastic for fast communication, it is not a tool designed for project management. It may work for projects with few co-authors but when the number of collaborators increase it becomes very difficult to keep track. Online systems for task management, if which most have mobile apps, include amongst otheres [Flow](http://www.getflow.com), [Asana](http://www.asana.com), [Wrike](http://www.wrike.com), [Basecamp](http://basecamp.com), Simplenote, [Evernote](www.evernote.com) and for Windows users there is OneNote.

**Step 6** We all learn by doing. When we create a reproducible workflow and share
reproduction materials we improve both cumulation of knowledge and our
methods for doing social science
[@freese2007replication,king1995replication].

# Research ought to be credible communication.

> [I]f the empirical basis for an article or book cannot be
  reproduced, of what use to the discipline are its
  conclusions?  What purpose does an article like this serve?
  [@king1995replication, p. 445]

We all always collaborate. Many of us collaborate with groups of people at one
moment in time as we race against a deadline. All of us collaborate with
ourselves over time. The time-frames over
which collaboration are required --- whether among a group of people working
together or within a single scholar's productive life or probably both --- are
much longer than any given version of any given software will easily exist.
Plain text is the exception.

But what if no one ever hears of your work, or, by some cruel fate, your
article does not spawn debate? Why then would you spend time to communicate
with your future self and others? Our own answer to this question is that it is efficient and that out of principle we
want our work to be credible and useful to ourselves and other scholars.  What we report in our data analyses should have two main
characteristics: (1) the findings of the work should not be a matter of
opinion; and (2) other people should be able to reproduce the findings. That
is, the work represents a shared experience --- and an experience shared
without respect to the identities of others (although requiring some common
technical training and research resources).

Assume we want others to believe us when we say something. More narrowly,
assume we want other people to believe us when we say something about data:
'data' here can be words, numbers, musical notes, images, ideas, etc. The point
is that we are making some claims about patterns in some collection of stuff.
For example, when Jake was invited into the homes and offices of ordinary
people in Chile in 1991, the "stuff" was recordings of conversations that he
had about life during the first year of democracy. Now, it might be easy to
convince others that 'this collection of stuff' is different from 'that
collection of stuff' if those people were looking over our shoulders the whole
time that we made decisions about collecting the stuff and broke it up into
understandable parts and reorganized and summarized it.  Unfortunately, we
can't assume that people are willing to shadow a researcher throughout her
career. Rather, we do our work alone or in small groups and want to convince
other distant and future people to take our analyses and findings seriously.

Now, say your collections of stuff are large or complex and your
chosen tools of analyses are computer programs. How can we convince
people that what we did with some data with some code is credible,
not a matter of whim or opinion, and reproducible by others who didn't
shadow us as we wrote our papers? This essay has suggested a few
concrete ways to enhance the believability of such scholarly work. In
addition, these actions (as summarized in the section headings of this
essay) make collaboration within research groups more effective.
Believability comes in part from reproducibility and researchers
often need to be able to reproduce in part or in whole what different
people in the group have done or what they, themselves, did in the past.

In the end, following these practices and those recommended by
@fredrickson2011tpm and @healy2011tpm among others working on these topics
allows your computerized analyses of your collections of stuff to be credible.
If then someone quibbles with your analyses, your future self can shoot them
the archive required to reproduce your work.^[Since you used plain text, the
files will still be intelligible, analyzed using commented code so that folks
can translate to whatever system succeeds R, or since you used R, you can
include a copy of R and all of the R packages you used in your final analyses
in the archive itself. You can even throw in a copy of whatever version of
Linux you used and an open source virtual machine running the whole environment
using say, Docker.] You can say, 'Here is everything you need to reproduce my
work.' To be extra helpful you can add 'Read the README file for further
instructions.' And then you can get on with your life: maybe the next great
idea will occur when your 4-year-old asks a wacky question after stripping and
painting her overly cooperative 1-year-old brother purple, or teaching a class,
or in a coffee shop, or on a quiet walk.

#References